Summer 2025 ToDo

### 2025-11-06 Relations/Paper/Well-foundedness.tex
Commenting out the old WF graph

The logical relationships between all of the above definitions are summarized in Figure \ref{fig:WF}.
All of the implications in the figure have been
formalized within the WellFounded subdirectory of the Agda repository.
An arrow from one node to another with no label indicates direct logical implication,
provable constructively with no additional assumptions. An arrow with a label indicates an additional
(semi-)classical hypothesis needed to establish the implication.
The definitions of these hypotheses are given in Figure \ref{tab:cprop}.

@@

\begin{figure}[h]
\centering
\begin{tikzpicture}[
  scale=0.7,
  every node/.style={transform shape},
  node distance=1cm and .5cm,
  box/.style={draw, rectangle, rounded corners, minimum width=1cm, minimum height=1cm, align=center},
  arrow/.style={-{Latex}, thick}
]

% Row 1: WFmin -> WFmin-
\node[box] (WFmin) {$\WFmin$};
\node[box, right=2.9cm of WFmin] (WFminM) {$\WFminm$};

\draw[arrow] (WFmin) -- (WFminM);

% Row 2: WFminDNE -> WFminDNE-
\node[box, below=of WFmin] (WFminDNE) {$\WFminDNE$};
\node[box, below=of WFminM] (WFminDNEM) {$\WFminDNEm$};

\draw[arrow] (WFminDNE) -- (WFminDNEM);

% Row 3: WFacc

\node[box, line width = 0.5mm, below=of WFminDNE] (WFacc) {$\WFacc$};

\node[box, below = of WFminDNEM] (WFaccM) {$\WFaccm$};
\draw[arrow] (WFacc) to (WFaccM);
\draw[arrow, bend left] (WFaccM) to node[below, text=red] {$\accDNE$} (WFacc);

% Row 4: WFseq
\node[box, below=of WFacc] (WFseq) {$\WFseq$};
\node[box, below= of WFaccM] (WFseqM) {$\WFseqm$};

\draw[arrow] (WFseq) -- (WFseqM);

% Row 5: WFcor
\node[box, below=of WFseq] (WFCor) {$\WFcor$};
\node[box, below=of WFseqM] (WFminCor) {$\WFcorm$};

% \draw[arrow, bend left = 55] (WFaccM) to (WFminM);

% Inter-row arrows
\draw[arrow] (WFacc) to node[right, text = red] {$\rdec$} (WFseq);
\draw[arrow, bend right] (WFaccM) to (WFminDNEM);
\draw[arrow, bend right = 90] (WFminCor) to node[pos = 0.4, right, text=red] {$\accCor$} (WFaccM);

\draw[arrow, bend right = 75] (WFmin) to (WFseq);
\draw[arrow, bend left = 75] (WFminM) to (WFseqM);
\draw[arrow] (WFaccM) -- (WFseqM);
\draw[arrow] (WFmin) -- (WFminDNE);
\draw[arrow, bend right] (WFminM) to (WFminDNEM);
\draw[arrow, bend right] (WFminDNEM) to (WFminM);

\draw[arrow] (WFminDNE) to node[right, text = red] {$\accDNE$} (WFacc);
\draw[arrow, bend right=45] (WFminDNE) to node[pos=0.3, left, text=red] {$\mpe$} (WFseq);
\draw[arrow, bend right] (WFminDNEM) to node[left, text=red] {$\FB$} (WFaccM);
\draw[arrow, bend right] (WFseqM) to (WFminCor);
\draw[arrow, bend right] (WFminCor) to node [right, text=red] {$\mpe$} (WFseqM);
\draw[arrow] (WFCor) -- (WFminCor);
% \draw[arrow, bend left = 110] (WFCor) to node[left, text=red] {$\accCor$} (WFacc);
\draw[arrow, bend left] (WFminCor) to node[below, text=red] {$\corDNE$} (WFCor);

\draw[arrow, bend left=80] (WFCor) to node[pos = 0.4, left, text=red] {$\accCor$} (WFacc);

\end{tikzpicture}
\caption{The logical relationships between notions of well-foundedness}
 \label{fig:WF}
\end{figure}
% \clearpage

@@

\end{tikzpicture}
\caption{The logical relationships between notions of well-foundedness}
\label{fig:WF2}
\end{figure}

@@

{
\def\arraystretch{1.3}
\begin{figure}[h!]
\small
\begin{tabular}{@{}l l l @{}}
\toprule
\textbf{Property} & \textbf{Definition}  &\textbf{Agda Code}  \\
\midrule
$\rdec$   & $R$ is decidable
          & $\forall$ \verb|{x} {y}| $\to$ \verb|EM (R x y)|\\
$\FB$     & $R$ is finitely branching
          & $\mathtt{\forall (a : A) \to \Sigma [ xs \in List A ] (\forall b \to R a b \to b \in{}List\, xs)}$ \\
$\mpe$    & Sequences are $\lnot\lnot$-closed
          & $\forall\ \mathtt{(f : \nat \to A) \to \lnot\lnot Closed
          (\lambda x \to \Sigma [ k \in \nat ] (f\, k \equiv x))}$ \\
$\corDNE$ & Coreductives are $\lnot\lnot$-closed
          & $\mathtt{\forall (P : \mathcal{P} A) \to R}$-$\mathtt{coreductive\ P \to \lnot\lnot Closed\ P}$\\
$\accDNE$ & Accessibility is $\lnot\lnot$-closed
          & $\lnot\lnot$\verb|Closed (R -accessible)| \\
$\accCor$ & Accessibility is coreductive
          & \verb|R -coreductive (R -accessible)| \\
\bottomrule
\end{tabular}
\centering
\caption{Classical Properties Used in the Well-Foundedness Diagram}
\label{tab:cprop}
\end{figure}
}
\clearpage
% The principles appearing in the labels of Figure \ref{fig:WF} are valid for many rewriting systems,
% and are classically true.
Figure \ref{fig:WF} illustrates that the network of logical relationships between
constructive notions of well-foundedness is quite rich indeed.
Taking the $\lnot\lnot$-closures of
these notions cuts down some, but not all, of the complexity.
For example, the two minimality principles become equivalent.
% The implication from the inductive to sequential definition no longer requires
% decidability of $R$.
% On the other hand, the reverse
At the same time, the weak version of the sequential definition is unable to imply
the (weak) accessible one without an additional hypothesis $\accCor$,
which may be described as the constructive contrapositive of the definition of accessibility.
In the context of rewriting systems,
this condition asserts the existence of a function which maps any
element that is \emph{not} strongly normalizing to a one-step reduct of it having the same property.
Assuming $\accCor$, all notions of well-foundedness
are constructively equivalent up to $\lnot\lnot$.

@@

\apcomment{Todo: 1. Filter through the rest, and see if there's anything we may still want to keep.\\
  2. Include a subsection, or a section, with examples.  Provide examples where some decidability
properties are provable, and others are not.}

>>>>

The logical relationships between all of the above definitions are summarized in Figure~\ref{fig:WF}.
All of the implications in the figure have been
formalized within the WellFounded subdirectory of the Agda repository.
An arrow from one node to another with no label indicates direct logical implication,
provable constructively with no additional assumptions. An arrow with a label indicates an additional
(semi-)classical hypothesis needed to establish the implication.
The definitions of these hypotheses are given in Figure \ref{tab:cprop}.

@@

% \begin{figure}[h]
% \centering
% \begin{tikzpicture}[
%   scale=0.7,
%   every node/.style={transform shape},
%   node distance=1cm and .5cm,
%   box/.style={draw, rectangle, rounded corners, minimum width=1cm, minimum height=1cm, align=center},
%   arrow/.style={-{Latex}, thick}
% ]

% % Row 1: WFmin -> WFmin-
% \node[box] (WFmin) {$\WFmin$};
% \node[box, right=2.9cm of WFmin] (WFminM) {$\WFminm$};

% \draw[arrow] (WFmin) -- (WFminM);

% % Row 2: WFminDNE -> WFminDNE-
% \node[box, below=of WFmin] (WFminDNE) {$\WFminDNE$};
% \node[box, below=of WFminM] (WFminDNEM) {$\WFminDNEm$};

% \draw[arrow] (WFminDNE) -- (WFminDNEM);

% % Row 3: WFacc

% \node[box, line width = 0.5mm, below=of WFminDNE] (WFacc) {$\WFacc$};

% \node[box, below = of WFminDNEM] (WFaccM) {$\WFaccm$};
% \draw[arrow] (WFacc) to (WFaccM);
% \draw[arrow, bend left] (WFaccM) to node[below, text=red] {$\accDNE$} (WFacc);

% % Row 4: WFseq
% \node[box, below=of WFacc] (WFseq) {$\WFseq$};
% \node[box, below= of WFaccM] (WFseqM) {$\WFseqm$};

% \draw[arrow] (WFseq) -- (WFseqM);

% % Row 5: WFcor
% \node[box, below=of WFseq] (WFCor) {$\WFcor$};
% \node[box, below=of WFseqM] (WFminCor) {$\WFcorm$};

% % \draw[arrow, bend left = 55] (WFaccM) to (WFminM);

% % Inter-row arrows
% \draw[arrow] (WFacc) to node[right, text = red] {$\rdec$} (WFseq);
% \draw[arrow, bend right] (WFaccM) to (WFminDNEM);
% \draw[arrow, bend right = 90] (WFminCor) to node[pos = 0.4, right, text=red] {$\accCor$} (WFaccM);

% \draw[arrow, bend right = 75] (WFmin) to (WFseq);
% \draw[arrow, bend left = 75] (WFminM) to (WFseqM);
% \draw[arrow] (WFaccM) -- (WFseqM);
% \draw[arrow] (WFmin) -- (WFminDNE);
% \draw[arrow, bend right] (WFminM) to (WFminDNEM);
% \draw[arrow, bend right] (WFminDNEM) to (WFminM);

% \draw[arrow] (WFminDNE) to node[right, text = red] {$\accDNE$} (WFacc);
% \draw[arrow, bend right=45] (WFminDNE) to node[pos=0.3, left, text=red] {$\mpe$} (WFseq);
% \draw[arrow, bend right] (WFminDNEM) to node[left, text=red] {$\FB$} (WFaccM);
% \draw[arrow, bend right] (WFseqM) to (WFminCor);
% \draw[arrow, bend right] (WFminCor) to node [right, text=red] {$\mpe$} (WFseqM);
% \draw[arrow] (WFCor) -- (WFminCor);
% % \draw[arrow, bend left = 110] (WFCor) to node[left, text=red] {$\accCor$} (WFacc);
% \draw[arrow, bend left] (WFminCor) to node[below, text=red] {$\corDNE$} (WFCor);

% \draw[arrow, bend left=80] (WFCor) to node[pos = 0.4, left, text=red] {$\accCor$} (WFacc);

% \end{tikzpicture}
% \caption{The logical relationships between notions of well-foundedness}
%  \label{fig:WF}
% \end{figure}
% % \clearpage

@@

\end{tikzpicture}
\caption{The logical relationships between notions of well-foundedness}
\label{fig:WF}
\end{figure}

@@

{
\def\arraystretch{1.3}
\begin{figure}[h!]
\small
\begin{tabular}{@{}l l l @{}}
\toprule
\textbf{Property} & \textbf{Definition}  &\textbf{Agda Code}  \\
\midrule
$\rdec$   & $R$ is decidable
          & $\forall$ \verb|{x} {y}| $\to$ \verb|EM (R x y)|\\
$\FB$     & $R$ is finitely branching
          & $\mathtt{\forall (a : A) \to \Sigma [ xs \in List A ] (\forall b \to R a b \to b \in{}List\, xs)}$ \\
$\mpe$    & Sequences are $\lnot\lnot$-closed
          & $\forall\ \mathtt{(f : \nat \to A) \to \lnot\lnot Closed
          (\lambda x \to \Sigma [ k \in \nat ] (f\, k \equiv x))}$ \\
$\corDNE$ & Coreductives are $\lnot\lnot$-closed
          & $\mathtt{\forall (P : \mathcal{P} A) \to R}$-$\mathtt{coreductive\ P \to \lnot\lnot Closed\ P}$\\
$\accDNE$ & Accessibility is $\lnot\lnot$-closed
          & $\lnot\lnot$\verb|Closed (R -accessible)| \\
$\accCor$ & Accessibility is coreductive
          & \verb|R -coreductive (R -accessible)| \\
\bottomrule
\end{tabular}
\centering
\caption{Classical Properties Used in the Well-Foundedness Diagram}
\label{tab:cprop}
\end{figure}
}
% \clearpage
% The principles appearing in the labels of Figure \ref{fig:WF} are valid for many rewriting systems,
% and are classically true.
Figure \ref{fig:WF} illustrates that the network of logical relationships between
constructive notions of well-foundedness is quite rich indeed.
Taking the $\lnot\lnot$-closures of
these notions cuts down some, but not all, of the complexity.
For example, the two minimality principles become equivalent.
% The implication from the inductive to sequential definition no longer requires
% decidability of $R$.
% On the other hand, the reverse
At the same time, the weak version of the sequential definition is unable to imply
the (weak) accessible one without an additional hypothesis $\accCor$,
which may be described as the constructive contrapositive of the definition of accessibility.
In the context of rewriting systems,
this condition asserts the existence of a function which maps any
element that is \emph{not} strongly normalizing to a one-step reduct of it having the same property.
Assuming $\accCor$, all notions of well-foundedness
are constructively equivalent up to $\lnot\lnot$.

@@

% \apcomment{Todo: 1. Filter through the rest, and see if there's anything we may still want to keep.\\
%   2. Include a subsection, or a section, with examples.  Provide examples where some decidability
% properties are provable, and others are not.}
